---
title: "R Notebook"
output: html_notebook
---

#Univariates 

```{r}
location="https://github.com/adamh27/Visual-Data/raw/main/"
file="Median%20Income.csv"

linkToFile=paste0(location,file)
```

```{r}
library(foreign)

median_income=read.csv(linkToFile)
```

```{r}
str(median_income)
```

```{r}
library(foreign) #unknown library 

median_income=read.csv(linkToFile) #load command to read data set in R

```
```{r}
str(median_income) #view data frame to ensure proper import 
```
```{r}
complete.cases(median_income) #Identify the NA values with T/F verification 

#2520 Omitted entries
```
```{r}
which(complete.cases(median_income)) #Shows which row values have complete values
```
```{r}
which(!complete.cases(median_income)) #Which rows have the NA values
```
```{r}
na_vec <- which(!complete.cases(median_income)) #Storing the NA variables here
na_vec #View the new na_vec
```
```{r}
No_NA_Median_Income <- median_income[-na_vec,] #Removing the NA data from the original data set and creating a new data set as N0_Na...

#1,491 obs removed 
```

```{r}
No_NA_Median_Income$Data <- as.numeric(as.character(No_NA_Median_Income$Data)) 
```

```{r}
str(No_NA_Median_Income) #Checking if numeric change occured
```

```{r}
#now I want to transform the data from a pure count into percents
head(No_NA_Median_Income$Race,20)
```
```{r}
#we want to build a table and transform the data to get percents

absoluteT=table(No_NA_Median_Income$Race,
                exclude = "nothing")

absoluteT
```

```{r}
prop.table(absoluteT)
```
```{r}
#want to transform the data to be a traditional percentage value
ToPlot=prop.table(absoluteT)*100

ToPlot
```

```{r}
# transform it into a data frame

tableFreq=as.data.frame(ToPlot)

names(tableFreq)=c("Race","pct")

tableFreq
```

```{r}
library(ggplot2)

```
```{r}
Abase= ggplot(data = tableFreq, 
             aes(x = Race,
                 y = pct)) 
```

```{r}
#Creating bar chart using the base instructions created in the previous chunk
plotA1 = Abase + geom_bar(fill ="grey",
                        stat = 'identity') 
#using the color black to make the chart readable

# I need this to see the result:
plotA1

```

```{r}
#Creating titles and labels for the the chart
titleText="Percent of Each Racial Group"
sourceText='Source: American Community Survey'

plotA2 = plotA1 + labs(title=titleText,
                     x =NULL, 
                     y = NULL,
                     caption = sourceText)
plotA2
```

```{r}
#changing the theme to make the chart more readable
plotA3=plotA2+ theme_classic() + coord_flip()+ theme(axis.text = element_text(size = 10))
plotA3
```

```{r}
#Want to change the limits/breaks to put the individual bars in perspective
library(scales) # for "unit_format""

# customize Y axis
plotA4 = plotA3 + scale_y_continuous(breaks=c(0,10, 20),
                                   limits = c(0, 20), 
                                   labels=unit_format(suffix = '%')) 
plotA4

```

```{r}
plotA5 = plotA4 + theme(plot.caption = element_text(hjust = 0), 
                      plot.title = element_text(hjust = 0.5))
plotA5

```


```{r}
#I want to add a line in the chart at the average, to see which racial groups are under and over represented in the data set 

plotA6 = plotA5 + geom_hline(yintercept = 14, #the average
                           linetype="dashed", 
                           size=1.5, #thickness
                           alpha=0.5) #transparency
plotA6
```

```{r}
#I want to tilt the racial titles to make the chart redable in a non expanded format

```

```{r}
#Lastly, I want to add the percentages for each bar to allow for better comparisons

LABELS=paste0(round(tableFreq$pct,2), '%')

plotA7 = plotA6 + geom_text(vjust=0,#adjust the height
                            hjust=2.5,
                          size = 4,#font face = "bold",
                          aes(y = pct ,
                              label = LABELS))
plotA7 #final non-interactive plot


```


```{r}
library(plotly) #make it interactive 

ggplotly(plotA7)
```

UNIVARIATE TWO
```{r}
na_vec <- which(!complete.cases(median_income))
```
```{r}
NA_median_income <- median_income[-na_vec,]
```
```{r}
NA_median_income$Data <- as.numeric(as.character(NA_median_income$Data))
```
```{r}
NA_median_income=NA_median_income[complete.cases(NA_median_income$Data),]
```

```{r}
aggregate(data=NA_median_income, Data~Location,median)

```


```{r}
str(NA_median_income)
```




```{r}
library(ggplot2)

base=ggplot(data=NA_median_income, # data frame
            aes(x=Data)) # what to plot 
```
```{r}
#Creating histogram chart using the base instructions created in the previous chunk
plot1= base + geom_histogram()
# I need this to see the result:
plot1 
```
```{r}
# all text for titles
Titles=list(Title="Washington State",
            SubTi="Median income",
            XTi="Median Income",
            YTi="Observations",
            Sou="Source: ACS")
#
# using the previous texts
plot2= plot1 + labs(title = Titles$Title,
                    subtitle = Titles$SubTi,
                    x=Titles$XTi,
                    y=Titles$YTi,
                    caption = Titles$Sou)
plot2
```
```{r}
#texts for annotation
message="Univariate"

#annotating
plot3=plot2 + annotate(geom = 'text',
                       x =150000,y=200, #position
                       label=message)
plot3
```
```{r}
plot4=plot3+ theme_minimal() #background change
plot4
```
```{r}
library(plotly)

ggplotly(plot4)
```
After Aes make interactive - see session 3 for uni numerical, histogram shape, bins, etc.

add a legend w SD, median, mean and make interactive with bins 

BIVARIATE 

```{r}
#Now we are interested in understanding how income differs across racial groups, this will be a CAT- NUM Bivariate visualization 

names(No_NA_Median_Income)
```

```{r}
summary(No_NA_Median_Income$Data)
```

```{r}
# looking at how out of place the max is, looking at the top 20 earners. 

library(magrittr) # for %>%
No_NA_Median_Income[order(-No_NA_Median_Income$Data),c('Race','Data')]%>%head(20)
```

```{r}
#Lets get the average of each group
aggregate(data=No_NA_Median_Income, Data~Race,median)
```

```{r}
#A box plot is an excellent way to visualize income data between the different groups 
library(ggplot2)
box=ggplot(data=No_NA_Median_Income,
            aes(x=Race,
                y=Data))

box + geom_boxplot() 
```

BIVARIATE TWO 

```{r}
NA_median_income=NA_median_income[complete.cases(NA_median_income$Data),]
```
```{r}
summary(NA_median_income)
```
```{r}
aggregate(data=NA_median_income, Data~Location,median) #combining multiple observations to get median for each county 
```


```{r}
# boxplot of median income by county 

library(ggplot2)
base=ggplot(data=simple_vec,
            aes(x= reorder(Location, -Data),y=Data)) #order bars in descending value 

plot1.1 = base + geom_bar(stat = "identity", width = 0.5, fill="steelblue" )
```

```{r}
# all text for titles
Titles=list(Title="Washington State",
            SubTi="County Level Median Income",
            XTi="County",
            YTi="Aggregate Median Income",
            Sou="Source: ACS")
#
# using the previous texts
plot2.2= plot1.1 + labs(title = Titles$Title,
                    subtitle = Titles$SubTi,
                    x=Titles$XTi,
                    y=Titles$YTi,
                    caption = Titles$Sou)
plot2.2
```
```{r}
plot3.3=plot2.2+ theme_minimal()
plot3.3 #change the theme of the plot 
```
```{r}
plot4.4=plot3.3+ coord_flip()+ theme(axis.text = element_text(size = 5))
plot4.4 #rotate plot and change variable text size 
```
```{r}
library(plotly) #making interactive 

ggplotly(plot4.4)
```
Rename columns to show "Data" as median income in the form of currency and remove the "reorder" command line from the interactive location naming 

MAPS 

```{r}
library(sf)
library(ggplot2)
library(tmap)
library(tmaptools)
library(leaflet)
library(dplyr)
```

```{r}
NA_median_income <- median_income[-na_vec,]
```
```{r}
NA_median_income$Data <- as.numeric(as.character(NA_median_income$Data)) 
```

```{r}
NA_median_income=NA_median_income[complete.cases(NA_median_income$Data),]
```
```{r}
simple_vec <- aggregate(data=NA_median_income, Data~Location,median)
simple_vec #creating a data frame with to have only one observation per county 
```

```{r}
str(NA_median_income)
```

```{r}
MapLocation="https://github.com/JMurray2015/hello-world/raw/main/"
MapFile="county10.json"

LinkMapFile=paste0(MapLocation,MapFile)

```
```{r}
CountyMap <- read_sf(LinkMapFile, stringsAsFactors= FALSE)
```
```{r}
str(CountyMap)
```
```{r}
colnames(CountyMap) #viewing column names of map data 
```
```{r}
names(CountyMap)[names(CountyMap) == "NAMELSAD10"] <- "Location"
```
```{r}
head(CountyMap)
```

```{r}
MergeMap=merge(CountyMap, #map first
                   simple_vec, 
                   by='Location')  
```
```{r}
str(MergeMap) #checking to see how the merge went 
```
```{r}
library(ggplot2)
# plot original map
Mapbase=ggplot(data=CountyMap) + geom_sf(fill='grey90',
                                     color=NA) + theme_classic()
Mapbase

```

Dashboard 

---
title: "543 Dashboard"
output:
    flexdashboard::flex_dashboard: 
      orientation: rows
      social: menu
      source: embed
     df_print: paged
---
# Intro Text {.sidebar}

The following information was sourced from the American Community Survey and includes the racial and ethnic distribution as well as median income within each Washington State county. Please find aggregate information on page 1, county level information on page 2, and county level mapping on page 3. 


```{r setup, include=FALSE}
library(flexdashboard)
knitr::opts_chunk$set(error = TRUE)
```

# Univariate Plots 
## R Markdown
### box 


```{r} 
library(plotly)

ggplotly(plot4)
```
Additional work: illustrate the sample mean, SD and confidence intervals - change observations to frequency 

## Including Plots
### box
```{r}
library(plotly)

ggplotly(plotA7)
```

# Bivariate Plots
## Column - column names dont appear in dashboard 
### box 

```{r pressure, echo=FALSE}
library(plotly)

ggplotly(plot4.4)
```
### box 

```{r}
library(plotly)

ggplotly(box2)
```
Additional work: Either flip coord or adjust group names  

# Maps
## idk 
### box 
```{r}
library(ggplot2)
# plot original map
base=ggplot(data=CountyMap) + geom_sf(fill='grey90',
                                     color=NA) + theme_classic()
base
```

Questions***: I am having a lot of difficulty merging the csv data with the map data despite renaming columns and the str/summary post merge appearing to have been correct. Have attempted shapefile and json, the methods in the session notes as well as multiple externally sourced. 

Are we able to load the maps in manually via (ggmap), map_data, tmap functions such as tm_shape, tm_borders, etc. or another function? 

### box 
```{r}
map 
```
 
